---
title: "Exercicio 5.2"
author: "Bruno Barbosa"
date: "1/11/2019"
output: html_document
---

```{r setup, include=TRUE, message=FALSE}
# clean old variables
rm(list=ls())


# Load Packages
library(tseries)
library(zoo)
library(QRM)
library(mvtnorm) # for sampling from a multivariate normal or t distribution
library(ggplot2)
library(rugarch)
library(tibble)

```

### 1) Aumente a amostra (inclua mais duas acoes e acrescente os dados mais recentes) 

```{r data}

# Determine the start date and end date of our samples.
dataIni = "2000-01-01"
dataFim = "2018-12-31"


# Fetch Microsoft Data (msft)
msft <- tseries::get.hist.quote(instrument = "msft",
                                    start = dataIni,
                                    end = dataFim,
                                    quote = c("Adjusted", "Vol"))
  
# Fetch IBM Data (ibm)
ibm <- tseries::get.hist.quote(instrument = "ibm",
                                start = dataIni,
                                end = dataFim,
                                quote = c("Adjusted", "Vol"))
  
# Fetch Coke Data (coke)
coke <- tseries::get.hist.quote(instrument = "coke",
                                  start = dataIni,
                                  end = dataFim,
                                  quote = c("Adjusted", "Vol"))
  
# Fetch Walmart Data (WMT)
walmart <- tseries::get.hist.quote(instrument = "WMT",
                                 start = dataIni,
                                 end = dataFim,
                                 quote = c("Adjusted", "Vol"))
  
# Calculation the return of series
r_msft = zoo::coredata(diff(log(msft$Adjusted)))
r_ibm  = zoo::coredata(diff(log(ibm$Adjusted)))
r_coke = zoo::coredata(diff(log(coke$Adjusted)))
r_wall = zoo::coredata(diff(log(walmart$Adjusted)))

# merge returns in one matrix
return_data = cbind(r_msft, r_ibm, r_coke, r_wall)

# Removing variables that will no longer be used 
rm(list = c("dataIni", "dataFim"))
rm(list = c("msft", "ibm", "coke", "walmart"))
rm(list = c("r_msft", "r_ibm", "r_coke", "r_wall"))

# Drop the first 14 observations
return_data = return_data[-c(1:14),]

# Calculate new sample size
T = nrow(return_data)

# Descriptive Summary of Data
summary(return_data)

```

### 2) Calcule o V@R(1%) de um portifólio com essas 4 acoes.
```{r portifolioSetup}
# Set Portfolio value
value = 1000

# V@R probability
p = 0.01 

# p% smallest
op = T*p


# Set Portfolio weights
w = matrix(c(0.25, 0.25, 0.25, 0.25), nrow = 4, ncol = 1)
  
# Portfolio returns
r_port = return_data %*% w

# Portfolio volatility
sigma_port = sqrt(t(w) %*% cov(return_data) %*% w)

  
```

##### V@R assuming Historical Simulation
```{r Var_HS}
# Sort the portifolio
r_port.sorted = sort(r_port)

# Portifolio V@R (Historical Simulation)
VaR_port.HS = -r_port.sorted[op] * value
print(VaR_port.HS)
  
# Univariate Expected shortfall for Portfolio
ES_port.HS = - mean(r_port.sorted[1:op]) * value
print(ES_port.HS)
```

##### V@R assuming Normal distribution
```{r Var_Normal}
# compute $\sigma$*$F^{-1}(p)$*Value
VaR_port.norm = -sigma_port *qnorm(p)*value
VaR_port.norm = as.numeric(VaR_port.norm)
print(VaR_port.norm)
```

##### V@R assuming Student-t distribution for the portifolio returns
obs: first we need to estimate the degrees of freedom of the distribution
```{r Var_t_stud}
# scale the returns
sc_r_port = r_port *100
  
# estimate the distribution parameters
res_port = QRM::fit.st(sc_r_port)
  
# rescale the volatility
sigma_sc_r_port = res_port$par.ests["sigma"]/100
  
# extract the degrees of freedom
nu_sc_r_port = res_port$par.ests[1]
  
# calculates the V@R
VaR_port.tstd = -sigma_sc_r_port * qt(df=nu_sc_r_port, p=p)*value
print(VaR_port.tstd)
  
# Remove auxiliary variables
rm(list = c("res_port","sc_r_port", "sigma_sc_r_port", "nu_sc_r_port", "r_port.sorted"))
```

##### V@R assuming multivariate Student-t distribution for the portifolio assets
```{r Var_t_stud_mult}
# Fitting a multivariate Student-t distribution
fit <- QRM::fit.mst(return_data, method = "BFGS")

# Portfolio multivariate Student-t V@R
sigma_port.M_tstd = sqrt(t(w) %*% fit$covariance %*%w)
VaR_port.M_tstd = - sigma_port.M_tstd * qt(p=p, df=fit$df) * value
VaR_port.M_tstd = as.numeric(VaR_port.M_tstd)
print(VaR_port.M_tstd)
  
# Remove unused variables
rm(list = c("fit", "sigma_port.M_tstd"))
```

##### Histogram of log returns of the portifolio and Normal distribution
```{r Histogram}
# density curve
d <- density(r_port)

# return table and density table (needed for ggplot)
tb_port = tibble(ret = r_port)
tb_port.density = tibble(x = d$x, y=d$y)

# Histogram plot
ggplot(data=tb_port, aes(tb_port$ret)) + 
  geom_histogram(aes(y=..density..), 
                 fill="black", 
                 alpha = .5,
                 bins = 50) +
  geom_line(aes(y = dnorm(tb_port$ret, mean(tb_port$ret), sd(tb_port$ret)),
                colour = "Normal"), size = 0.8) +
  geom_line(data = tb_port.density,
            aes(x = x, y = y, colour = "Kernel"),
            size = 0.8) +
  scale_colour_manual(values=c("red","blue")) +
  labs(title= "Histogram", # Title
       subtitle="Portifolio Histogram and Normal distribution", # Subtitle
       caption=NULL, # Caption
       y="Density", 
       x='Retuns',
       color=NULL)

# remove unused variables
rm(list = c("d", "tb_port", "tb_port.density"))
```

##### Expected shortfall for Portfolio with Normal distribution
```{r Expected shortfall normal}
# Expected shortfall for Portfolio with Normal distribution
ES_port.normal = sigma_port*dnorm(qnorm(p))/p*value
ES_port.normal = as.numeric(ES_port.normal)
print(ES_port.normal)
```

##### Moving Average V@R for portfolio 
```{r Moving Average}
# Initialize the moving average object
VaR_port.MA <- rep(NA, length(r_port))

# RiskMetrics: lambda = 0.94 and r = 0
lambda=0.94;

# Window: 20 days
WE=20
for (t in seq(WE+1,length(r_port))){
  t1=t-WE+1
  window = r_port[t1:t] # estimation window
  sigma=sd(window)
  VaR.MA = -sigma * qnorm(p) * value
  VaR_port.MA[t]=VaR.MA
}

# remove unused variables
rm(list = c("WE", "t", "t1", "window", "sigma", "VaR.MA"))

# MA normal VaR and VaR(1%)

# Create a table with VaR_port.MA and VaR_port.norm
VaR_port.table = tibble(day = 1:length(VaR_port.MA), VaR_MA = VaR_port.MA, VaR_Norm = VaR_port.norm)

# Plot the graphic
ggplot(VaR_port.table) + 
  geom_line(aes(x=day, y=VaR_MA, colour="MA(20)")) +
  geom_line(aes(x=day, y=VaR_Norm, colour="Normal"), linetype = "dashed") +
  scale_colour_manual(values=c("black","red")) +
  labs(title= "V@R for the Portifolio", # Title
       subtitle="V@R-MA(20) and V@R-Normal(1%)", # Subtitle
       caption=NULL, # Caption
       y=NULL, 
       x=NULL,
       color=NULL)
```

##### Exponential weighted moving average model VaR for Portfolio 
```{r EWMA}
# Estiates a Exponential weighted moving average model VaR for Portfolio 

# Initialize the EWMA object
VaR_port.EWMA <- rep(NA,length(r_port))

# Covariance matrix of the portifolio
s = cov(return_data)

for (t in seq(2,length(r_port))){
  s=lambda*s+(1-lambda)*return_data[t-1,]%*%t(return_data[t-1,])
  sigma=sqrt(t(w)%*%s%*%w)
  VaR_port.EWMA[t]=-sigma*qnorm(p)*value
}

# remove unused variables
rm(list = c("t", "s", "sigma"))

# Add the V@R estimatives to the table
VaR_port.table$VaR_EWMA = VaR_port.EWMA

ggplot(VaR_port.table) + 
  geom_line(aes(x=day, y=VaR_EWMA, colour="EWMA")) +
  geom_line(aes(x=day, y=VaR_Norm, colour="Normal"), linetype = "dashed") +
  scale_colour_manual(values=c("black","red")) +
  labs(title= "V@R for the Portifolio", # Title
       subtitle="V@R-EWMA and V@R-Normal(1%)", # Subtitle
       caption=NULL, # Caption
       y=NULL, 
       x=NULL,
       color=NULL)
```

### Passo 3) Experimente vários modelos garch, mas escolha um. Justifique a escolha

```{r garch}
# Especifica ARCH(1) com Distribuicao Normal
spec_arch = rugarch::ugarchspec(variance.model=list(model="sGARCH",garchOrder = c(1, 0)),
                                mean.model=list(armaOrder=c(0,0)),
                                distribution.model = "norm")

# Especifica GARCH(1,1) com Distribuicao Normal
spec_garch = rugarch::ugarchspec(variance.model = list(model="sGARCH", garchOrder = c(1, 1)),
                                 mean.model = list( armaOrder = c(0,0), include.mean = FALSE),
                                 distribution.model = "norm")


# Especifica EGARCH(1,1) com Distribuicao Normal
spec_egarch = ugarchspec(variance.model=list(model="eGARCH",garchOrder = c(1, 1)),
                        mean.model=list(armaOrder=c(0,0)),
                        distribution.model = "norm")


# Especifica GjrGARCH(1,1) com Distribuicao Normal
spec_gjrgarch = ugarchspec(variance.model=list(model="gjrGARCH", garchOrder = c(1, 1)),
                   mean.model=list(armaOrder=c(0,0)),
                   distribution.model = "norm")

# Fit the specified models
fit_arch = ugarchfit(spec = spec_arch, data = r_port)
fit_garch = ugarchfit(spec = spec_garch, data = r_port)
fit_egarch = ugarchfit(spec = spec_egarch, data = r_port)
fit_gjr = ugarchfit(spec = spec_gjrgarch, data = r_port)


# Determines the information criteria of the models
rugarch::infocriteria(fit_arch)
rugarch::infocriteria(fit_garch)
rugarch::infocriteria(fit_egarch)
rugarch::infocriteria(fit_gjr)
```

We must choose the best model. Although eGarch and Garch models can't be directly compared due to having different dependent variables on the volatility equation, we will disregard this and use the Bayes Information criteria for model selection)

* arch      - Bayes        -6.103389
* garch     - Bayes        -6.288601
* e-garch   - Bayes        -6.305992
* Gjr-garch - Bayes        -6.303335

By the BIC we choose the e-garch model.

```{r var garch}
# Copute the volatility of the distributions
fit_arch.sigma = rugarch::sigma(fit_arch)
fit_garch.sigma = rugarch::sigma(fit_garch)
fit_egarch.sigma = rugarch::sigma(fit_egarch)
fit_gjr.sigma = rugarch::sigma(fit_gjr)

# Initialize a vector of VAR for each garch model
VaR_port.arch <- rep(NA,length(r_port))
VaR_port.garch <- rep(NA,length(r_port))
VaR_port.egarch <- rep(NA,length(r_port))
VaR_port.gjrgarch <- rep(NA,length(r_port))

# Compute the VaR for each Garch model
for (t in seq(2,length(r_port))){
  VaR_port.arch[t]     = -fit_arch.sigma[t]*qnorm(p)*value
  VaR_port.garch[t]    = -fit_garch.sigma[t]*qnorm(p)*value
  VaR_port.egarch[t]   = -fit_egarch.sigma[t]*qnorm(p)*value
  VaR_port.gjrgarch[t] = -fit_gjr.sigma[t]*qnorm(p)*value
}

# Add the V@R estimatives for the Garch models to the table
VaR_port.table$VaR_ARCH = VaR_port.arch
VaR_port.table$VaR_GARCH = VaR_port.garch
VaR_port.table$VaR_EGARCH = VaR_port.egarch
VaR_port.table$VaR_GjrGARCH = VaR_port.gjrgarch

# Plot the graphic for each V@R estimative
ggplot(VaR_port.table) + 
  geom_line(aes(x=day, y=VaR_ARCH, colour="ARCH")) +
  geom_line(aes(x=day, y=VaR_Norm, colour="Normal"), linetype = "dashed") +
  scale_colour_manual(values=c("black","red")) +
  labs(title= "V@R for the Portifolio", # Title
       subtitle="V@R ARCH and V@R-Normal(1%)", # Subtitle
       caption=NULL, # Caption
       y=NULL, 
       x=NULL,
       color=NULL)


ggplot(VaR_port.table) + 
  geom_line(aes(x=day, y=VaR_GARCH, colour="GARCH")) +
  geom_line(aes(x=day, y=VaR_Norm, colour="Normal"), linetype = "dashed") +
  scale_colour_manual(values=c("black","red")) +
  labs(title= "V@R for the Portifolio", # Title
       subtitle="V@R GARCH and V@R-Normal(1%)", # Subtitle
       caption=NULL, # Caption
       y=NULL, 
       x=NULL,
       color=NULL)

ggplot(VaR_port.table) + 
  geom_line(aes(x=day, y=VaR_EGARCH, colour="e-GARCH")) +
  geom_line(aes(x=day, y=VaR_Norm, colour="Normal"), linetype = "dashed") +
  scale_colour_manual(values=c("black","red")) +
  labs(title= "V@R for the Portifolio", # Title
       subtitle="V@R e-GARCH and V@R-Normal(1%)", # Subtitle
       caption=NULL, # Caption
       y=NULL, 
       x=NULL,
       color=NULL)


ggplot(VaR_port.table) + 
  geom_line(aes(x=day, y=VaR_GjrGARCH, colour="Gjr-GARCH")) +
  geom_line(aes(x=day, y=VaR_Norm, colour="Normal"), linetype = "dashed") +
  scale_colour_manual(values=c("black","red")) +
  labs(title= "V@R for the Portifolio", # Title
       subtitle="V@R Gjr-GARCH and V@R-Normal(1%)", # Subtitle
       caption=NULL, # Caption
       y=NULL, 
       x=NULL,
       color=NULL)
```
